---
title: "prediction_cours8_project"
author: "Diyana Nanova"
date: "20/02/2022"
output: pdf_document
  #html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, include=FALSE}
library(lattice)
library(ggplot2)
library(caret)
library(rpart)
library(rpart.plot)
library(corrplot)
library(rattle)
library(randomForest)
library(RColorBrewer)
library(tibble)
library(bitops)
```


# Executive Summary 
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively.
The goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants and to apply a machine learning algorithm to the 20 test cases available in the test data and to submit the predictions.

# Data Source
The training and test data for this project are collected using the link below:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

# Exercise
## 1.Loading of Data
```{r message=FALSE}
url_train <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
url_test  <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"

data_train <- read.csv(url(url_train), strip.white = TRUE, na.strings = c("NA",""))
data_test  <- read.csv(url(url_test),  strip.white = TRUE, na.strings = c("NA",""))

dim(data_train)
dim(data_test)
```



## 2.Cleaning of Data
```{r message=FALSE}
#Remove firsdt 6 columns
data_train1 <- data_train[ , -(1:6)]
data_test1  <- data_test[ , -(1:6)]

#Remove variables that are mostly NA
data_train2 <- colnames(data_train1)[!colSums(is.na(data_train1)) > 0]
train_set <- data_train[data_train2]
data_test2 <- colnames(data_test1)[!colSums(is.na(data_test1)) > 0]
test_set <- data_test[data_test2]
```


## 3.Split Data into random train and test
```{r message=FALSE}
train_part = createDataPartition(train_set$classe, p = 0.75)[[1]]
training = train_set[ train_part,]
testing = train_set[-train_part,]

dim(training)
dim(testing)
```

## 4The different modells

### 4.1.Decision Tree Model
#### 4.1.1. Set the model
```{r message=FALSE}
set.seed(12345)
decision_tree_training <- rpart(classe ~ ., data = training, method="class")
rpart.plot(decision_tree_training)
```

#### 4.1.2.Predictions of the decision tree model on testing
```{r message=FALSE}
decision_tree_predicting <- predict(decision_tree_training, newdata = testing, type="class")
decision_tree_matrix <- confusionMatrix(decision_tree_predicting, factor(testing$classe))
decision_tree_matrix
```

### 4.2. Random Forest Model
#### 4.2.1. Set the model
```{r message=FALSE}
set.seed(12345)
rfm_control<- trainControl(method = "repeatedcv", number = 5, repeats = 2)
rfm_training <- train(classe ~ ., data = training, method = "rf",
                      trControl = rfm_control, verbose = FALSE)
```

#### 4.2.2.Predictions of the random forest model
```{r message=FALSE}
rfm_predicting <- predict(rfm_training, testing)
matrix_rfm <- confusionMatrix(rfm_predicting, factor(testing$classe))
matrix_rfm
```

### 4.3.Generalized Boosted Model (GBM)
#### 4.1.1. Set the model
```{r message=FALSE}
set.seed(12345)
gbm_control<- trainControl(method = "repeatedcv", number = 5, repeats = 2)
gbm_training <- train(classe ~ ., data = training, method = "gbm",
                      trControl = gbm_control, verbose = FALSE)
```

#### 4.1.2.Predictions of the generalized boosted model
```{r message=FALSE}
gbm_predicting <- predict(gbm_training, newdata = testing)
matrix_gbm <- confusionMatrix(gbm_predicting, factor(testing$classe))
matrix_gbm
```


## 5.Applying the Best Predictive Model to the Test Data

The following are the predictive accuracy of the three models:
  
Decision Tree Model: 74,14%
Generalized Boosted Model:  98,67%
Random Forest Model:  99,63%

The Random Forest Model is with a better accuracy and will be used to make predictions on the 20 data points from the original testing dataset(data_test).

```{r message=FALSE}
predict_modell <- as.data.frame(predict(rfm_training, newdata = data_test))
predict_modell
```

